Medical Concept Extraction

### 2012-12-17 paper/paper.tex

\subsubsection{MIT Regular Expressions}
In order to capture some general characteristics of each word we included a set of regular expressions each of which adds a dimension to our classification. Such features help not only to pick up on traits within different medical categories--such as tests which are often abbreviated or contain numbers--but also to easily rule out non-medical terms such as days, phone numbers, etc. The characteristics we chose are as follows:
\begin{itemize}
\item {\tt INITCAP}: first letter is capitalized and alphabetic,
\item {\tt ALLCAPS}: entire word is capitalized and alphabetic,
\item {\tt CAPSMIX}: the word is a mix of capital and lowercase alphabetic characters,
\item {\tt HASDIGIT}: there is a digit in the word,
\item {\tt SINGLEDIGIT}: the token is a single digit,
\item {\tt DOUBLEDIGIT}: the token is a two digits,
\item {\tt FOURDIGITS}: the token is a series of four digits,
\item {\tt NATURALNUM}: the token is a natural number,
\item {\tt REALNUM}: the token is a real number,
\item {\tt ALPHANUM}: the token contains only digits and alphabetic characters,
\item {\tt HASDASH}: the token has a dash,
\item {\tt PUNCTUATION}: the token is punctuation,
\item {\tt PHONE1}: the token is a phone number,
\item {\tt PHONE2}: the token is a phone number with area code,
\item {\tt FIVEDIGIT}: the token is a series of five digits,
\item {\tt NOVOWELS}: the token does not contain vowels (i.e. abbreviations),
\item {\tt HASDASHNUMALPHA}: the token consists of a dash, numbers, and alphabetic characters, 
\item {\tt DATESEPERATOR}: the token is a date separator (e.g. {\tt [-/]})
\end{itemize}

@@

\item Grid search and grid search image

\item Include that PNG image from easy.py

\item Cross validation

\section{System}

>>>>

\subsubsection{Regular Expressions}
In order to capture some general characteristics of each word we included a set of regular expressions each of which adds a dimension to our classification. Such features help allow not only to pick up on traits within different medical categories--such as tests which are often abbreviated or contain numbers--but also to easily rule out non-medical terms such as days, phone numbers, etc. The characteristics we chose are as follows:
\begin{itemize}
\item {\tt INITCAP}: first letter is capitalized and alphabetic,
\item {\tt ALLCAPS}: entire word is capitalized and alphabetic,
\item {\tt CAPSMIX}: the word is a mix of capital and lowercase alphabetic characters,
\item {\tt HASDIGIT}: there is a digit in the word,
\item {\tt SINGLEDIGIT}: the token is a single digit,
\item {\tt DOUBLEDIGIT}: the token is a two digits,
\item {\tt FOURDIGITS}: the token is a series of four digits,
\item {\tt NATURALNUM}: the token is a natural number,
\item {\tt REALNUM}: the token is a real number,
\item {\tt ALPHANUM}: the token contains only digits and alphabetic characters,
\item {\tt HASDASH}: the token has a dash,
\item {\tt PUNCTUATION}: the token is punctuation,
\item {\tt PHONE1}: the token is a phone number,
\item {\tt PHONE2}: the token is a phone number with area code,
\item {\tt FIVEDIGIT}: the token is a series of five digits,
\item {\tt NOVOWELS}: the token does not contain vowels (i.e. abbreviations),
\item {\tt HASDASHNUMALPHA}: the token consists of a dash, numbers, and alphabetic characters, 
\item {\tt DATESEPERATOR}: the token is a date separator (e.g. {\tt [-/]})
\end{itemize}

@@

Due to the data's bias with respect to labels which we previously discussed, the default SVM parameters failed to provide any sort of relevant results. Therefore, with our full feature set we used libsvm's included utilities to perform parameter selection as shown in Figure~\ref{fig:parameter_selection}. The utility performed $5$-fold cross validation and allowed us to see the parameters which provided the best accuracy. In turn, this enabled us to find an appropriate constant $C$ with which to scale the loss function and constant $\gamma$ for the radial basis function kernel the SVM was using.

It should be noted that the best method would be to optimize against the F1 score rather than the accuracy due to the data bias previously discussed. However, in practice, the parameters that were discovered using this method were able to dramatically improve performance while still being easy to obtain.

\section{System}
This section discusses the framework code we have developed in order to perform concept extraction, the web service frontend which allows users to test the service without seeing any of the underlying data, and the machine learning libraries that are leveraged.
